\large{\bf Abnormal quantitative genetics? }

### 2023-07-03 writeup.tex

Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
\[
	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
We then have
\begin{multline*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	= 
	\begin{cases}
		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
		 \begin{multlined}
		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
		\right)    
		\end{multlined}
		& \text{if $\alpha = 1$.}
	\end{cases}
\end{multline*}
This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha > 1$, in which case the limit is 
\[
	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\]
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

@@

Therefore,
to study how this works in practice, we turn to simulation.
We used SLiM v4 \citep{haller2022slim4}
to simulate a discrete-time approximation to the continuous-time Moran model described above
as follows.
Individuals are diploid, hermaphroditic, and sexual;
the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
mutations occur at a rate of $m$ per bp.
Each new mutation is independently assigned an ``effect''
drawn from the ``effect size distribution'' (and will be either Gaussian or Cauchy).
The phenotype of each individual is determined additively
by summing the effects of all mutations they carry across both chromosomes.
Let $f(z) = 1 + \mu(z)$
be the ``fitness'' of an individual with phenotype $z$,
and let $dt$ be a small constant (we took $dt=0.01$).
Then, at each time step
each individual dies with probability $1 - \exp(-dt f(z))$;
if there are $k$ such individuals chosen to die in a given time step,
then we choose $k$ individuals uniformly and without replacement
to produce new offspring;
for each such reproduction event the parent chooses a mate uniformly at random from the population.
This maintains the population at a fixed size, $N$.
(Note that individuals may self and individuals chosen to die may also reproduce,
but in a large population these details should be unimportant.)

@@

>>>>

Now, consider the case when $\Psi(\theta)$ corresponds to a stable law, so
\[
	\Psi(\theta) = im\theta + c|\theta|^{\alpha} (1-i\beta \text{sgn}(\theta)\omega_{\alpha}(\theta))
\]
for $\alpha \in (0,2]$, $\beta \in [-1,1]$, $c \geq 0$, and $m \in \mathbb{R}$ and
\[
	\omega_{\alpha}(\theta) = \begin{cases}
		\tan{\frac{\pi\alpha}{2}} & \text{if $\alpha \neq 1$, and}\\
		-\frac{2}{\pi}\ln{|\theta|} & \text{if $\alpha = 1$.}
	\end{cases}
\]
We then have
\begin{multline*}
	 \sum_{k=1}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}
	 =  \sum_{k=1}^{n-1} im\theta + 2^{(1-\alpha)k} c|\theta|^{\alpha} {\textstyle\left(1-i \beta\text{sgn}(\theta)\left(\frac{\theta}{2^{k}}\right)
	 	 \omega_{\alpha}\left(\frac{\theta}{2^{k}}\right)\right)}\\
	= 
	\begin{cases}
		 im(n-1)\theta + c|\theta|^{\alpha}\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right) & \text{if $\alpha \neq 1$, and}\\	
		 \begin{multlined}
		 im(n-1)\theta + c|\theta|^{\alpha}  \left(\frac{1-2^{(1-\alpha)n}}{1-2^{1-\alpha}}
		 \left(1+i \beta \text{sgn}(\theta)\frac{2}{\pi}\ln{|\theta|} \right)\right.\\
		\left. -i\beta \text{sgn}(\theta) \frac{2^{1-\alpha} + (n(2^{1-\alpha}-1)-1)2^{(1-\alpha)n})}{(1-2^{1-\alpha})^{2}}\frac{2}{\pi}\ln{2}
		\right)    
		\end{multlined}
		& \text{if $\alpha = 1$.}
	\end{cases}
\end{multline*}
This converges as $n \to \infty$ if and only if $m = 0$ and $\alpha \ge 1$, in which case the limit is 
\[
	 \frac{c}{1-2^{1-\alpha}}|\theta|^{\alpha}\left(1-i\beta \text{sgn}(\theta) \tan{\frac{\pi\alpha}{2}}\right), 
\]
which corresponds to a $\left(\alpha,\beta, \frac{c}{1-2^{1-\alpha}},0\right)$-stable law.

@@

Therefore,
to study how this works in practice, we turn to simulation.
We used SLiM v4 \citep{haller2022slim4}
to simulate a discrete-time approximation to the continuous-time Moran model described above
as follows.
Individuals are diploid, hermaphroditic, and sexual;
the genome is of length $10^8$bp with a uniform recombination rate of $10^{-8}$ crossovers per bp;
mutations occur at a rate of $10^{-9}$ per bp.
Each new mutation is independently assigned an ``effect''
drawn from the ``effect size distribution'' (and will be either Gaussian or Cauchy).
The phenotype of each individual is determined additively
by summing the effects of all mutations they carry across both chromosomes.
Let $f(z) = 1 + \mu(z)$
be the ``fitness'' of an individual with phenotype $z$,
and let $dt$ be a small constant (we took $dt=0.01$).
Then, at each time step
each individual dies with probability $1 - \exp(-dt f(z))$;
if there are $k$ such individuals chosen to die in a given time step,
then we choose $k$ individuals uniformly and without replacement
to produce new offspring;
for each such reproduction event the parent chooses a mate uniformly at random from the population.
This maintains the population at a fixed size, $N$.
(Note that individuals may self and individuals chosen to die may also reproduce,
but in a large population these details should be unimportant.)

@@

\subsection{Notes}

Let $x_\ell$ and $p_\ell$ be sequences of numbers.
When is, for given $s_1$, $s_2$, $t_1$, and $t_2$,
\begin{align*}
    &
    \prod_{\ell=1}^M
    \left\{
    \frac{1}{2}
        \left( 1 + p_\ell(e^{i s_1 x_\ell/M^{1/\alpha}} - 1) \right)
        \left( 1 + p_\ell(e^{i s_2 x_\ell/M^{1/\alpha}} - 1) \right)
        +
    \frac{1}{2}
        \left( 1 + p_\ell(e^{i t_1 x_\ell/M^{1/\alpha}} - 1) \right)
        \left( 1 + p_\ell(e^{i t_2 x_\ell/M^{1/\alpha}} - 1) \right)
    \right\}
\\ & \qquad \approx
    \prod_{\ell=1}^M
        \left( 1 + p_\ell (e^{i s_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
        \left( 1 + p_\ell (e^{i s_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
        \left( 1 + p_\ell (e^{i t_1 x_\ell/M^{1/\alpha}} - 1)/2 \right)
        \left( 1 + p_\ell (e^{i t_2 x_\ell/M^{1/\alpha}} - 1)/2 \right)
\end{align*}
